---
title: "God in Philosophy"
output: html_notebook
---

# 1. Introduction

In the beginning of *History of Western Philosophy*, Russell gives a beautiful introduction to the subject Philosophy by defining it as a middle ground between Science and Theology:    

>“Philosophy, as I shall understand the word, is something intermediate between theology and science. Like theology, it consists of speculations on matters as to which deﬁnite knowledge has, so far, been unascertainable; but like science, it appeals to human reason rather than to authority, whether that of tradition or that of revelation. All deﬁnite knowledge—so I should contend—belongs to science; all dogma as to what surpasses deﬁnite knowledge belongs to theology. But between theology and science there is a No Man’s Land, exposed to attack from both sides; this No Man’s Land is philosophy.” 

The concept of God is an important topic of many philosophy articles. Philosophers have sought to justify it or to demolish it or to show that it lies outside reason's grasp. By analyzing the history of philosophy data set, we are going to explore the sentiments of different philosophy schools towards God and the main topics when they discussed about God.

# 2. Data Processing

```{r echo=FALSE, message=FALSE, warning=FALSE}
packages.used=c("rvest", "tibble", "qdap", 
                "sentimentr", "gplots", "dplyr",
                "tm", "syuzhet", "factoextra", 
                "beeswarm", "scales", "RColorBrewer",
                "RANN", "tm", "topicmodels","readtext","wordcloud","stringr")

# check packages that need to be installed.
packages.needed=setdiff(packages.used, 
                        intersect(installed.packages()[,1], 
                                  packages.used))
# install additional packages
if(length(packages.needed)>0){
  install.packages(packages.needed, dependencies = TRUE)
}

# load packages
library(dplyr)
library(tidyr)
library("rvest")
library("tibble")
# You may need to run
# sudo ln -f -s $(/usr/libexec/java_home)/jre/lib/server/libjvm.dylib /usr/local/lib
# in order to load qdap
library("qdap")
library("sentimentr")
library("gplots")
library("dplyr")
library("tm")
library("syuzhet")
library("factoextra")
library("beeswarm")
library("scales")
library("RColorBrewer")
library("RANN")
library("tm")
library("topicmodels")
library("readtext")
library(forcats)
library(ggplot2)
library(beeswarm)
library(wordcloud)
library(stringr)
```

The History of Philosophy Dataset is available for download at https://www.kaggle.com/kouroshalizadeh/history-of-philosophy. It is a .csv file that includes 360808 observations and 11 variables. These observations are sentences taken from 59 texts and 13 different schools spanning the history of philosophy. Below shows some samples of observations and a list of all 13 schools involved.

```{r echo=FALSE, message=FALSE, warning=FALSE}
philosophy.data  <- read.csv("/Users/xilezhang/OneDrive/GitHub/philosophy_data.csv")
head(philosophy.data)
unique(philosophy.data$school)
```
In this project, we mainly focus on theology topic in Philosophy. Therefore, we filter sentences containing keywords *god* or *lord* to finish following analysis. The filtered dataset contains 8183 sentences.

```{r echo=FALSE, message=FALSE, warning=FALSE}
god.data = rbind(philosophy.data[str_detect(philosophy.data$tokenized_txt,'god'),],
                 philosophy.data[str_detect(philosophy.data$tokenized_txt,'lord'),])
dim(god.data)
```

# 3. Data Analysis: Sentence Distribution

In the following, I want to check how these sentences distribute over years and different schools. First two charts below show that school Rationalism spent great portion of texts discussing topics related to God, so as Stoicism, Plato and Nietzsche. In addition, in 350 B.C. and between 1650 A.C. and 1750 B.C., theology may be a popular topic.  

```{r echo=FALSE, message=FALSE, warning=FALSE}

total.sentence.number <- philosophy.data %>%
  group_by(school) %>%
  summarise(total = n())
god.sentence.number <- god.data %>%
  group_by(school) %>%
  summarise(sub = n()) 
total.sentence.number %>%
  left_join(god.sentence.number, by="school") %>%
  mutate(rate = sub/total) %>%
  ggplot(aes(x = school, y = rate))+
  geom_bar(stat = "identity", fill="#f68060", alpha=.6, width=.4)+
  coord_flip()+
  theme_bw()+
  scale_y_continuous(labels = percent) +
  labs(title = "Ratio of sentences directly referencing God over different schools",
       x = "school", y = "number of sentences")


god.data %>%
  group_by(original_publication_date)%>%
  summarise(num_sentence = n())%>%
  ggplot(aes(x=original_publication_date, y=num_sentence)) +
  #geom_area(fill="#69b3a2", alpha=0.5) +
  geom_line(color="#69b3a2") +
  ylab("Number of sentences") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))+
  scale_x_continuous(breaks = seq(-350,1985,200)) +
  labs(title = "Number of sentences related to God over years")


```
What's more, below graph indicates different schools devote various length to the description of God. Rationalism describe this field in details, and so do Plato and Empiricism. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

beeswarm(sentence_length~school, data = god.data, horizontal=TRUE,
         pch=16, col=alpha(brewer.pal(9, "Set1"), 0.6), 
         cex=0.55, cex.axis=0.8, cex.lab=0.8,
         spacing=1.2/nlevels(factor(god.data$school)),
         las=2, xlab="Number of Words in a Sentence.", ylab="",
         main="The distribution of sentence lengths related to God over different school")

```

# 4. Data Analysis: Sentiment Analysis

## 4.1 Text Proprocessing

Here method Sentiment Analysis is used to show authors' emotion behind the sentences. Below are some examples of results of sentiment analysis.

```{r echo=FALSE, message=FALSE, warning=FALSE}
sentence.list=NULL
for(i in 1:nrow(god.data)){
  sentences=god.data$sentence_lowered[i]
  if(length(sentences)>0){
    emotions=get_nrc_sentiment(sentences)
    word.count=word_count(sentences)
    emotions=(1/(word.count+0.01))%*%as.matrix(emotions)
    sentence.list=rbind(sentence.list, 
                        cbind(god.data[i,-ncol(god.data)],
                              sentences=as.character(sentences), 
                              word.count,
                              emotions,
                              sent.id=1:length(sentences)
                              )
    )
  }
}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
head(sentence.list)
```

Below are some emotionally charged sentences of several representative schools, Rationalism, Plato and Nietzsche. It is kind of obvious that the represented sentences of schools Rationalism and Plato is pretty positive towards the concept of God. Especially Rationalism related God with many virtues like knowledge, truth and perfection, which suggesting rationality and religion are not completely antagonistic. In fact, Rationalism has a branch called theistic rationalism, which believes natural religion, Christianity, and rationalism typically coexist compatibly. On the contrary, represented sentences of school Nietzsche and Feminism is relatively negative, which is also proved in the table below. 
```{r echo=FALSE, message=FALSE, warning=FALSE}
print("Rationalism")
speech.df=tbl_df(sentence.list)%>%
  filter(school=="rationalism")%>%
  select(sentences, anger:trust)
speech.df=as.data.frame(speech.df)
as.character(speech.df$sentences[apply(speech.df[,-1], 2, which.max)])

print("Plato")
speech.df=tbl_df(sentence.list)%>%
  filter(school=="plato")%>%
  select(sentences, anger:trust)
speech.df=as.data.frame(speech.df)
as.character(speech.df$sentences[apply(speech.df[,-1], 2, which.max)])

print("Nietzsche")
speech.df=tbl_df(sentence.list)%>%
  filter(school=="nietzsche")%>%
  select(sentences, anger:trust)
speech.df=as.data.frame(speech.df)
as.character(speech.df$sentences[apply(speech.df[,-1], 2, which.max)])

print("Feminism")
speech.df=tbl_df(sentence.list)%>%
  filter(school=="feminism")%>%
  select(sentences, anger:trust)
speech.df=as.data.frame(speech.df)
as.character(speech.df$sentences[apply(speech.df[,-1], 2, which.max)])
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
sentence.list %>%
  group_by(school) %>%
  summarise_each(negative, funs = mean) %>%
  arrange(desc(negative))
```

## 4.2 Clustering of emotions
The result of heat map is consistent with common sense that sentiments of same type would have stronger connection with each other. The heat map and bar chart of Sentiment Distribution below also indicate that four protrude sentiments towards God: trust, anticipation, joy and fear; mostly positive.

```{r echo=FALSE, message=FALSE, warning=FALSE}
heatmap.2(cor(sentence.list%>%select(anger:trust)), 
          scale = "none", 
          col = bluered(100), margin=c(6, 6), key=F,
          trace = "none", density.info = "none")

par(mar=c(4, 6, 2, 1))
emo.means=colMeans(select(sentence.list, anger:trust)>0.01)
col.use=c("red2", "darkgoldenrod1", 
            "chartreuse3", "blueviolet",
            "darkgoldenrod2", "dodgerblue3", 
            "darkgoldenrod1", "darkgoldenrod1")
barplot(emo.means[order(emo.means)], las=2, col=col.use[order(emo.means)], horiz=T, main="Sentiment Distribution")
```

## 4.3 KMeans Clustering: grouping authors emotions

```{r echo=FALSE, message=FALSE, warning=FALSE}
set.seed(2021)

presid.summary=tbl_df(sentence.list)%>%
  #group_by(paste0(type, File))%>%
  group_by(author)%>%
  summarise(
    anger=mean(anger),
    anticipation=mean(anticipation),
    disgust=mean(disgust),
    fear=mean(fear),
    joy=mean(joy),
    sadness=mean(sadness),
    surprise=mean(surprise),
    trust=mean(trust)
    #negative=mean(negative),
    #positive=mean(positive)
  )
presid.summary=as.data.frame(presid.summary)
rownames(presid.summary)=as.character((presid.summary[,1]))
km.res=kmeans(presid.summary[,-1], iter.max=200,
              3)
fviz_cluster(km.res, 
             stand=F, repel= TRUE,
             data = presid.summary[,-1], xlab="", xaxt="n",
             show.clust.cent=FALSE)
```
From above clustering chart, we apply K-Means on sentiment analysis result to split authors into 3 different groups. Table below shows that not all authors from same school will fall into the same group, suggesting that philosophers from same school may have different attitude towards theology topic.

```{r echo=FALSE, message=FALSE, warning=FALSE}
school.author <- god.data %>%
  group_by(school, author) %>%
  summarise()
  
author <- as_tibble(names(km.res$cluster))
cluster <- as_tibble(km.res$cluster)
tibble(cluster,author,.name_repair = "unique") %>%
  rename(cluster=value...1, author=value...2) %>%
  left_join(school.author, by = "author") %>%
  arrange(cluster, school)
```

# 5. Data Analysis - Topic Modeling 
## 5.1 Topic Modeling 

Topic modeling is a type of statistical modeling for discovering the abstract “topics” that occur in a collection of documents. Here I am going to apply LDA to the dataset and split these sentences into topics. 

```{r echo=FALSE, message=FALSE, warning=FALSE}

sentence_corpus <- Corpus(VectorSource(god.data$sentence_lowered))
sentence_corpus <- tm_map(sentence_corpus, removeNumbers)
sentence_corpus <- tm_map(sentence_corpus, removePunctuation)
sentence_corpus <- tm_map(sentence_corpus, removeWords, c("the", "and", "also", "thus","therefore", stopwords("english")))
sentence_corpus <- tm_map(sentence_corpus, stripWhitespace)

dtm <- DocumentTermMatrix(sentence_corpus)

rowTotals <- apply(dtm , 1, sum) #Find the sum of words in each Document

#corpus.list <- corpus.list[rowTotals>0, ]
dtm  <- dtm[rowTotals> 0, ]
god.data <- god.data[rowTotals>0, ]

#Set parameters for Gibbs sampling
burnin <- 4000
iter <- 2000
thin <- 500
seed <-list(2003,5,63,100001,765)
nstart <- 5
best <- TRUE

#Number of topics
k <- 15
library(topicmodels)
#Run LDA using Gibbs sampling
ldaOut <-LDA(dtm, k, method="Gibbs", control=list(nstart=nstart, 
                                                 seed = seed, best=best,
                                                 burnin = burnin, iter = iter, 
                                                 thin=thin))


ldaOut.topics <- as.matrix(topics(ldaOut))
ldaOut.terms <- as.matrix(terms(ldaOut,20))
topicProbabilities <- as.data.frame(ldaOut@gamma)

```


```{r echo=FALSE, message=FALSE, warning=FALSE}
#Analysing topics from all schools
terms.beta=ldaOut@beta
terms.beta=scale(terms.beta)
topics.terms=NULL
for(i in 1:k){
  topics.terms=rbind(topics.terms, ldaOut@terms[order(terms.beta[i,], decreasing = TRUE)[1:7]])
}
topics.terms
```

Based on the most popular terms and the most salient terms for each topic, I manually assign a hashtag to each topic: "creator", "ruler", "history", "practice", "human", "sense", "appellation", "capital", "sacrifice", "prove", "impact", "location", "time", "salvation", "language". This assignment process is not objective and does not guarantee accuracy. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
topics.hash=c("creator", "ruler", "history", "practice", "human", "sense", "appellation", "capital", "sacrifice", "prove", "impact", "location", "time", "salvation", "language")
sentence.list$ldatopic=as.vector(ldaOut.topics)
sentence.list$ldahash=topics.hash[ldaOut.topics]

colnames(topicProbabilities)=topics.hash

corpus.list <- data.frame(text=sapply(sentence.list, identity), 
    stringsAsFactors=F)

femaile_all_t <- god.data[, c("title","author","school","original_publication_date","sentence_length","sentence_lowered")]
corpus.list.df <- cbind(femaile_all_t, topicProbabilities)
head(corpus.list.df)

```

## 5.2 Clustering of topics
```{r echo=FALSE, message=FALSE, warning=FALSE}
library(gplots)

topic.summary=tbl_df(corpus.list.df)%>%
              select(school, creator:language)%>%
              group_by(school)%>%
              summarise_each(funs(mean))

topic.summary=as.data.frame(topic.summary)
rownames(topic.summary)=topic.summary[,1]
#rownames(topic.summary) <- make.names(topic.summary[,1], unique=TRUE)

topic.plot=c(1, 2, 5, 8, 9, 11, 14)

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
heatmap.2(as.matrix(topic.summary[,topic.plot+1]), 
          scale = "column", key=F, 
          col = bluered(100),
          cexRow = 0.9, cexCol = 0.9, margins = c(8, 8),
          trace = "none", density.info = "none")

```

Above heatmap shows that Rationalism has positive correlation with identity attribute of God such as creator, ruler and savior. On the contrary, School Feminism and Communism more concentrate on the human topic instead of the attribute of God.

# 6. Data Analysis - Word Embedding

## 6.1 Word Embedding

Next, I would use word embedding method to obtain the closest words among the sentences for word *God*. It is not surprising that word *God* has great similarity with words like *man*, *himself* based on the patriarchy in ancient Greek and medieval.

``` {r echo=FALSE, message=FALSE, warning=FALSE}
library(word2vec)

x = philosophy.data$sentence_lowered
model <- word2vec(x = x, type = "cbow", dim = 15, iter = 20)
#embedding <- as.matrix(model)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
lookslike <- predict(model, "god" , type = "nearest", top_n = 10)
lookslike
```

Nietzsche was a philosopher who was extremely critical of religion. He believes that people should regard themselves as their own beliefs rather than belief in God, and once said a famous saying: God is dead. Therefore, as expected, *God* term was attached with some negative terms such as *devil*, *enemy* in his sentences.

```{r echo=FALSE, message=FALSE, warning=FALSE}
philosophy.data[str_detect(philosophy.data$sentence_lowered,'god is dead'),] %>%
  filter(author == "Nietzsche") %>%
  select(title, author, school, sentence_spacy)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
nietzsche <- philosophy.data$sentence_lowered[philosophy.data$school == "nietzsche"]
model.nietzsche <- word2vec(x = nietzsche, type = "cbow", dim = 15, iter = 20)
embedding.nietzsche <- as.matrix(model.nietzsche)
predict(model.nietzsche, "god" , type = "nearest", top_n = 10)
```

As analysed before, Rationalism and Plato have positive sentiment towards theology, and more focus on the identity of God. The close words of God in their sentences like divinity, creator, 	intellect, grace also reflects this tendency.
```{r echo=FALSE, message=FALSE, warning=FALSE}
rationalism <- philosophy.data$sentence_lowered[philosophy.data$school == "rationalism" | philosophy.data$school == "plato"]
model.rationalism <- word2vec(x = rationalism, type = "cbow", dim = 15, iter = 20)
#embedding <- as.matrix(model)
predict(model.rationalism, "god" , type = "nearest", top_n = 10)
```

## 6.2 Word Cloud

A word cloud graph will be drawed to intuitively see those keywords related to word *God*. 

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(syuzhet)
library(lubridate)
library(ggplot2)
library(scales)
library(reshape2)
library(dplyr)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
sentence.rationalism <- god.data$sentence_lowered
corpus.rationalism <- iconv(AP_text, to = 'utf-8-mac')
corpus.rationalism <- Corpus(VectorSource(corpus_AP))
corpus.rationalism <- tm_map(corpus_AP, tolower)
corpus.rationalism <- tm_map(corpus_AP, removePunctuation)
corpus.rationalism <- tm_map(corpus_AP, removeNumbers)
corpus.rationalism <- tm_map(corpus_AP, removeWords, stopwords('english'))
corpus.rationalism <- tm_map(corpus_AP, removeWords, c("god", "the", "but", "one", "also", "will",  "just","another", "either", "sometimes", "always", "gods"))
corpus.rationalism <- tm_map(corpus.rationalism, stripWhitespace)
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
dtm.rationalism <- TermDocumentMatrix(corpus.rationalism)
dtm.rationalism <- removeSparseTerms(dtm.rationalism, 0.99)
tidy.rationalism <- tidytext::tidy(dtm.rationalism)
summarise.rationalism <- summarise(group_by(tidy.rationalism, term), sum(count))

```

```{r echo=FALSE, message=FALSE, warning=FALSE}
wordcloud(summarise.rationalism$term, summarise.rationalism$`sum(count)`,
          scale=c(5,0.5),
          max.words=100,
          min.freq=1,
          random.order=FALSE,
          rot.per=0.3,
          random.color=FALSE,
          colors=brewer.pal(9,"Blues"))
```

From above word cloud, we can see that keywords such as nature, mind, power, love, order are frequently used by philosophers when they talked about God.


# 7. Conclusion
All in all, based on the analysis, we can conclude that different philosophy school had different opinion and attitude towards theology topics. School Rationalism and Plato spend a lot of sentences discussing theology topics and had a positive sentiment towards God. On the contrary, School Feminism and Nietzsche, though they also spent a lot of sentences discussing the concept of God, they held a relatively critical and negative sentiment towards God and may more focus on human itself. Overall, mostly philosophers showed mixed but largely positive sentiments towards the image of God and associated it with creator, love, power and nature.






